---
title: "Statistics 101A Project - Group 5: Pokemon Weights"
author: "Ryan Chu, Samvit Garg, Nicole Lee, Naren Prakash, Nathaniel Pranoto Tjong, Oliver Siu"
date: "2024-03-17"
output:
  html_document:
    toc: yes
    toc_depth: 4
    toc_float: yes
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 0. Sourcing

```{r source}
library(tidyverse)
library(car)
library(leaps)
pokemon <- read_csv("pokemon.csv", show_col_types = FALSE)
```

## 1. Original Dataset

```{r p1}
pokemon
dim(pokemon) #dataset dimensions

# non-numeric inputs detected
pokemon$capture_rate[is.na(as.numeric(pokemon$capture_rate))]

# modified dataset for compatibility
pokemod <- read_csv("pokemon_mod.csv", show_col_types = FALSE)
pokemod
```

## 2. Subsetted Dataset

### a) Subsetting

```{r p2a, cache = TRUE}
pokemon_subset <- pokemod[, c("weight_kg", "base_egg_steps", "attack", "defense", "hp",
                              "sp_attack", "sp_defense", "capture_rate", "experience_growth")]
pokemon_subset
pairs(pokemon_subset)
```

### b) Original Model

```{r p2b}
pokemon_full_model <- lm(weight_kg ~ capture_rate + base_egg_steps + attack + defense + hp +
                           sp_attack + sp_defense + capture_rate + experience_growth,
                         data = pokemon_subset)
summary(pokemon_full_model)
```

### c) NA values


```{r  p2c}
which(!complete.cases(pokemon_subset))
pokemon_subset[which(!complete.cases(pokemon_subset)), ]
```


### d) Original Model Diagnostics

```{r p2d, cache = TRUE}
par(mfrow = c(2, 2))
plot(pokemon_full_model)
```

### e) Original Model Formula

```{r p2e}
cof <- summary(pokemon_full_model)$coefficients
```

$$\begin{aligned}
\text{weight}
&= `r cof[1]`
+ `r cof[2]` * \text{base egg steps}
+ `r cof[3]` * \text{attack}
+ `r cof[4]` * \text{defense} \\
&+ `r cof[5]` * \text{hp}
+ `r cof[6]` * \text{sp_attack}
`r cof[7]` * \text{sp_defense} \\
&`r cof[8]` * \text{capture_rate}
+ `r cof[9]` * \text{experience_growth}
\end{aligned}$$

## 3. Box-cox Transformation on X & Y

### a) Power Transform on X & Y

```{r p3a}
pokemon_full_model_pt <- powerTransform(pokemon_subset)
summary(pokemon_full_model_pt)

attach(pokemon_subset)
t_weight <- weight_kg^0.16
t_bes <- base_egg_steps^-0.67
t_attack <- attack^0.5
t_defense <- defense^0.33
t_hp <- hp^0.33
t_sa <- sp_attack^0.33
t_sd <- sp_defense^0.15
t_cp <- capture_rate^0.33

t_pokemon_model <- lm(t_weight ~ t_bes + t_attack + t_defense + t_hp +
                        t_sa + t_sd + t_cp + experience_growth)
summary(t_pokemon_model)
```

### b) Full Box-Cox Diagnostics

```{r p3b, cache = TRUE}
par(mfrow = c(2, 2))
plot(t_pokemon_model)
```

### c) Box-Cox Model Formula

```{r p3c}
pow <- summary(pokemon_full_model_pt)$result[, 2]
cof <- summary(t_pokemon_model)$coefficients
```

Box-cox model:
$$\begin{aligned}
\text{weight}^{`r pow[1]`}
&= `r cof[1]`
`r cof[2]` * \text{base egg steps}^{`r pow[2]`}
+ `r cof[3]` * \text{attack}^{`r pow[3]`}
+ `r cof[4]` * \text{defense}^{`r pow[4]`} \\
&+ `r cof[5]` * \text{hp}^{`r pow[5]`}
`r cof[6]` * \text{sp_attack}^{`r pow[6]`}
+ `r cof[7]` * \text{sp_defense}^{`r pow[7]`} \\
&`r cof[8]` * \text{capture_rate}^{`r pow[8]`}
+ `r cof[9]` * \text{experience_growth}
\end{aligned}$$

## 4. Box-Cox on X and Inverse Response on Y

### a) Power Transform on X

```{r p4a}
pokemon_subset2 <- pokemod[, c("weight_kg", "base_egg_steps", "attack", "defense", "hp",
                               "sp_attack", "sp_defense", "capture_rate", "experience_growth")]
attach(pokemon_subset2)
exp_model <- powerTransform((pokemon_subset2[, c("base_egg_steps", "attack", "defense", "hp",
                                                 "sp_attack", "sp_defense", "capture_rate", "experience_growth")]))
summary(exp_model)

t2_bes <- base_egg_steps^-0.67
t2_attack <- attack^0.5
t2_defense <- defense^0.33
t2_hp <- hp^0.33
t2_sa <- sp_attack^0.33
t2_sd <- sp_defense^0.14
t2_cp <- capture_rate^0.33

test2 <- lm(weight_kg ~ t2_bes + t2_attack + t2_defense + t2_hp + t2_sa + t2_sd + t2_cp)
```

### b) Inverse Response on Y

```{r p4b, cache = TRUE}
par(mfrow = c(1,1))
inverseResponsePlot(test2)

t2_weight <- weight_kg^0.35

t2_pokemon_model <- lm(t2_weight ~ t2_bes + t2_attack + t2_defense + t2_hp +
                         t2_sa + t2_sd + t2_cp + experience_growth)
```

### c) Box-Cox on X and Inverse Response on Y Diagnostics

```{r p4c, cache = TRUE}
par(mfrow = c(2, 2))
plot(t2_pokemon_model)
```

### d) Box-Cox on X and Inverse Response on Y Formula

```{r p4d}
pow <- summary(exp_model)$result[, 2]
cof <- t2_pokemon_model$coefficients
```

Inverse response plot model:
$$\begin{aligned}
\text{weight}^{0.35}
&= `r cof[1]`
`r cof[2]` * \text{base egg steps}^{`r pow[1]`}
+ `r cof[3]` * \text{attack}^{`r pow[2]`}
+ `r cof[4]` * \text{defense}^{`r pow[3]`} \\
&+ `r cof[5]` * \text{hp}^{`r pow[4]`}
`r cof[6]` * \text{sp_attack}^{`r pow[5]`}
+ `r cof[7]` * \text{sp_defense}^{`r pow[6]`} \\
&`r cof[8]` * \text{capture_rate}^{`r pow[7]`}
+ `r cof[9]` * \text{experience_growth}
\end{aligned}$$

## 5. Transformation Comparison

### a) Box-Cox

```{r p5a1}
vif(t_pokemon_model)
summary(t_pokemon_model)
```

```{r p5a2}
n <- nrow(pokemon_subset)
p <- 8

t_std_resid <- rstandard(t_pokemon_model)
t_leverages <- hatvalues(t_pokemon_model)
t_cooks <- cooks.distance(t_pokemon_model)

t_outlier <- t_std_resid[abs(t_std_resid) > 2]
t_leverage_point <- t_leverages[t_leverages > ((2 * (p + 1)) / n)]
t_influential <- t_cooks[t_cooks > 4 / (n - 2)]

t_outlier_by_leverage <- t_std_resid[t_leverages > ((2 * (p + 1)) / n)]
t_bad_leverage <- t_outlier_by_leverage[abs(t_outlier_by_leverage) > 2]
t_bad_leverage
```

### b) Inverse Response

```{r p5b1}
vif(t2_pokemon_model)
summary(t2_pokemon_model)
```

```{r p5b2}
n <- nrow(pokemon_subset)
p <- 8

t2_std_resid <- rstandard(t2_pokemon_model)
t2_leverages <- hatvalues(t2_pokemon_model)
t2_cooks <- cooks.distance(t2_pokemon_model)

t2_outlier <- t_std_resid[abs(t2_std_resid) > 2]
t2_leverage_point <- t2_leverages[t2_leverages > ((2 * (p + 1)) / n)]
t2_influential <- t2_cooks[t_cooks > 4 / (n - 2)]

t2_outlier_by_leverage <- t2_std_resid[t2_leverages > ((2 * (p + 1)) / n)]
t2_bad_leverage <- t2_outlier_by_leverage[abs(t2_outlier_by_leverage) > 2]
t2_bad_leverage
```

### c) Conclusion

We will choose the full Box-Cox model.

## 6. Variable Selection

### a) VIF and Plots

```{r p6a, cache = TRUE}
vif(t_pokemon_model)
pairs(t_weight ~ t_bes + t_attack + t_defense + t_hp + t_sa + t_sd + t_cp + experience_growth)
```

### b) All Possible Subsets

#### Best Subset Models

```{r p6b1}
transformed_pokemon_subset <- data.frame(cbind(t_weight ~ t_bes + t_attack + t_defense +
                                                 t_hp + t_sa + t_sd + t_cp + experience_growth))
pokemon_model_subsets <- regsubsets(t_weight ~ t_bes + t_attack + t_defense +
                                      t_hp + t_sa + t_sd + t_cp + experience_growth,
                                    data = transformed_pokemon_subset)
summary(pokemon_model_subsets)
```

#### Diagnostics

```{r p6b2}
p1 <- lm(weight_kg ~ t_hp)
p2 <- lm(weight_kg ~ t_defense+t_hp)
p3 <- lm(weight_kg ~ t_attack+t_defense+t_hp)
p4 <- lm(weight_kg ~ t_bes+t_attack+t_defense+t_hp)
p5 <- lm(weight_kg ~ t_bes+t_attack+t_defense+t_hp+experience_growth)
p6 <- lm(weight_kg ~ t_bes+t_attack+t_defense+t_hp+t_cp+experience_growth)
p7 <- lm(weight_kg ~ t_bes+t_attack+t_defense+t_hp+t_sd+t_cp+experience_growth)
p8 <- lm(weight_kg ~ t_bes+t_attack+t_defense+t_hp+t_sa+t_sd+t_cp+experience_growth)

pokemon_model_list <- list(p1, p2, p3, p4, p5, p6, p7, p8)

sub_fit <- function(x, n) {
  size <- seq_along(x)
  radj2 <- numeric(0)
  aic <- numeric(0)
  aic_cor <- numeric(0)
  bic <- numeric(0)
  for (i in seq_along(x)) {
    radj2 <- c(radj2, summary(x[[i]])$adj.r.squared)
    aic <- c(aic, extractAIC(x[[i]])[2])
    aic_cor <- c(aic_cor, extractAIC(x[[i]])[2]+ (2 * (i + 2) * (i + 3)) / (n - i - 1))
    bic <- c(bic, extractAIC(x[[i]], k = log(n))[2])
  }
  data.frame(Size = size, Radj2 = radj2, AIC = aic, AICc = aic_cor, BIC = bic)
}

pokemon_gof <- sub_fit(pokemon_model_list, n = nrow(pokemon_subset))
pokemon_gof
```

#### Best Models

```{r p6b3}
data.frame(Criteria = colnames(pokemon_gof)[-1], Model = c(which.max(pokemon_gof$Radj2),
                                                   which.min(pokemon_gof$AIC),
                                                   which.min(pokemon_gof$AICc),
                                                   which.min(pokemon_gof$BIC)))
```

### c) Stepwise Regression

#### Backwards AIC

```{r p6c1}
backAIC <- step(p8, direction = "backward")
```

#### Backwards BIC

```{r p6c2}
backBIC <- step(p8, direction = "backward", k = log(nrow(pokemon_subset)))
```

#### Forwards AIC

```{r p6c3}
pokemon_base <- lm(t_weight ~ 1)
forwardAIC <- step(pokemon_base,
                   scope = list(lower = ~ 1,
                                upper = ~ t_bes + t_attack + t_defense +
                                  t_hp + t_sa + t_sd + t_cp + experience_growth),
                   direction = "forward")
```

#### Forwards BIC

```{r p6c4}
pokemon_base <- lm(t_weight ~ 1)
forwardAIC <- step(pokemon_base,
                   scope = list(lower = ~ 1,
                                upper = ~ t_bes + t_attack + t_defense +
                                  t_hp + t_sa + t_sd + t_cp + experience_growth),
                   direction = "forward",
                   k = log(nrow(pokemon_subset)))
```

#### Best Models

- Backwards AIC: weight_kg ~ t_bes + t_defense + t_hp + t_sa + experience_growth
- Backwards BIC: weight_kg ~ t_bes + t_defense + t_hp + experience_growth
- Forwards AIC: t_weight ~ t_hp + t_defense + t_attack + t_bes + experience_growth + t_sd + t_sa + t_cp
- Forwards BIC: t_weight ~ t_hp + t_defense + t_attack + t_bes

### d) Conclusion

We went with full model.

## 7. Final Model

